\chapter{Abstract}

\begin{english} %switch to English language rules
Hardware and software engineers are often confronted by many different
integrated development environments. That big amount of environments
exists because in most cases there is one development environment for
each programming language. Very often engineers have to develop their
hardware and software on different operating systems. That adds more
environments to their list. To avoid this big amount of development
environments, Emacs will be introduced as an alternative.\\\\ This
thesis describes how Emacs can be used for hardware and software
development. Emacs requires to be configured specifically to the needs
of hardware and software engineers. It will be described, how the
Emacs configuration works. The configuration in this thesis modifies
Emacs for the use as an integrated development environment for the
programming languages C, C++, Emacs Lisp, Python and VHDL. In the
configuration file, there are packages offered by the Emacs
community. The functionality and configuration of all packages are
described. There are some additional functions implemented to offer a
simpler and quicker usage for the functionality of these
packages. Some packages add graphical user interface elements to
Emacs. These packages should animate people for the use of Emacs, that
are more likely to develop on graphical environments. Emacs is
introduced from scratch, especially for people who are completely new
to Emacs.\\\\ The thesis describes some possible workflows for the
development with different programming languages. This is meant to
offer software and hardware developers an easier and quicker access to
Emacs. That way it should be easy for them, to include Emacs in their
daily workflow.
\end{english}

